---
title: Age at First Diagnosis
author: Jacob Simmering
format: 
  html:
    embed-resources: true
---


```{r pkg-load}
library(tidyverse)
library(survival)
```

Living in an MSA seeks to be consistently associated with lower negative effects
of having a diagnosis for one of the ALS risk factors. One possible cause
of this may be people living in MSAs being diagnosed with ALS earlier on in
their disease course. One way to look at this would be to check the age at 
first diagnosis of ALS. 

```{r}
model_data <- read_rds("/Shared/lss_jsimmeri/als/model_data.rds")

model_data <- model_data |>
  mutate(
    age = first_year - dobyr,
    female = sex == 1,
    stratum = as.numeric(as.factor(glue::glue("{first_year}-{last_year}-{dobyr}-{sex}")))
  ) |>
  mutate(n_sx = fall + op + muscle,
         any_sx = fall | op | muscle)
```

Living in an MSA is associated with an younger age at diagnosis:

```{r}
survfit(
  Surv(age, als) ~ lives_in_msa,
  data = model_data
)
```

Residents of MSAs are, on median, 85 years old when diagnosed compared to 88 
among their peers that are non-MSA resident. 

The median age of diagnosis decreases with the number of symptoms:

```{r}
survfit(
  Surv(age, als) ~ n_sx,
  data = model_data
)
```

The decrease is present in both residents and non-residents but the residents
consistently have younger ages (although the difference may be n/s at times).

```{r}
survfit(
  Surv(age, als) ~ n_sx + lives_in_msa,
  data = model_data
)
```
