---
title: Main Analysis
author: Jacob Simmering
format: 
  html:
    embed-resources: true
---

```{r pkg-load}
library(tidyverse)
library(survival)
library(tableone)
library(gtsummary)
```

# Data Loading and Processing
```{r}
if (Sys.info()["sysname"] == "Darwin") {
  model_data <- read_rds("/Volumes/lss_jsimmeri/als/model_data.rds")
} else {
  model_data <- read_rds("/Shared/lss_jsimmeri/als/model_data.rds")
}
```

```{r}
model_data <- model_data |>
  mutate(
    age = first_year - dobyr,
    female = sex == 1,
    stratum = as.numeric(as.factor(glue::glue("{first_year}-{last_year}-{dobyr}-{sex}")))
  ) |>
  mutate(any_sx = motor | bulbar,
         n_groups = motor + bulbar,
         n_bulbar = speech + swallowing,
         n_motor = strength + gait + involuntary_movement + pain + other + falls,
         n_total = speech + swallowing + strength + gait + involuntary_movement + pain + other + falls)
```

# Data Summary
First, the data summary by ALS status:

```{r}
all_vars <- c(
  "age", "female", 
  "first_year", "last_year",
  "rate_outpatient_visits", "intensity_outpatient_visits_dx",
  "rate_inpatient_stays", "intensity_inpatient_stays_los",
  "elix_CHF", "elix_Valvular", "elix_PHTN", "elix_PVD", "elix_HTN",
  "elix_HTNcx", "elix_Paralysis", "elix_NeuroOther", "elix_Pulmonary", "elix_DM",
  "elix_DMcx", "elix_Hypothyroid", "elix_Renal", "elix_Liver", "elix_PUD", 
  "elix_HIV", "elix_Lymphoma", "elix_Mets", "elix_Tumor", "elix_Rheumatic", 
  "elix_Coagulopathy", "elix_Obesity", "elix_WeightLoss", "elix_FluidsLytes", "elix_BloodLoss", 
  "elix_Anemia", "elix_Alcohol", "elix_Drugs", "elix_Psychoses", "elix_Depression", 
  "lives_in_msa",
  "any_sx", "motor", "bulbar",
  "n_groups", "n_bulbar", "n_motor", "n_total",
  "speech", "swallowing", 
  "strength", "gait", "involuntary_movement", "pain", "other", "falls")
factor_vars <- c("n_groups", "n_bulbar", "n_motor", "n_total")
nonnormal_vars <- c("age", "first_year", "last_year", "rate_outpatient_visits", 
  "intensity_outpatient_visits_dx", "rate_inpatient_stays", 
  "intensity_inpatient_stays_los")

CreateTableOne(
  data = model_data,
  vars = all_vars,
  factorVars = factor_vars,
  strata = c("als")
) |>
  print(nonnormal = nonnormal_vars, smd = TRUE)
```

And then by ALS and whether they live in an MSA:

```{r}
CreateTableOne(
  data = model_data,
  vars = all_vars,
  factorVars = factor_vars,
  strata = c("als", "lives_in_msa")
) |>
  print(nonnormal = nonnormal_vars, smd = TRUE)
```

# Define Models
I want to fit XXXX models: 

First, does having any of these codes increase the odds of getting an ALS 
diagnosis?

```{r}
any_fml_unadj <- als ~ any_sx * lives_in_msa + strata(stratum)

any_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   any_sx * lives_in_msa + strata(stratum)
```

Second, what is the independent effect of bulbar symptoms?

```{r}
any_bulbar_unadj <- als ~ bulbar * lives_in_msa + strata(stratum)

any_bulbar_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   bulbar * lives_in_msa + strata(stratum)
```

Third, what is the independent effect of motor symptoms?

```{r}
any_motor_unadj <- als ~ motor * lives_in_msa + strata(stratum)

any_motor_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   motor * lives_in_msa + strata(stratum)
```

Fourth, what is the effect of symptoms from 0, 1, or 2 affects systems?

```{r}
number_origin_unadj <- als ~ as.factor(I(bulbar + motor)) * lives_in_msa + strata(stratum)

number_origin_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   as.factor(I(bulbar + motor)) * lives_in_msa + strata(stratum)
```

Fifth, what is the effect of the symptom sub-class? This is actually many 
models...

```{r}
speech_fml_unadj <- als ~ speech * lives_in_msa + strata(stratum)

speech_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   speech * lives_in_msa + strata(stratum)

swallowing_fml_unadj <- als ~ swallowing * lives_in_msa + strata(stratum)

swallowing_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   swallowing * lives_in_msa + strata(stratum)

strength_fml_unadj <- als ~ strength * lives_in_msa + strata(stratum)

strength_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   strength * lives_in_msa + strata(stratum)

gait_fml_unadj <- als ~ gait * lives_in_msa + strata(stratum)

gait_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   gait * lives_in_msa + strata(stratum)

involuntary_movement_fml_unadj <- als ~ involuntary_movement * lives_in_msa + strata(stratum)

involuntary_movement_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   involuntary_movement * lives_in_msa + strata(stratum)

pain_fml_unadj <- als ~ pain * lives_in_msa + strata(stratum)

pain_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   pain * lives_in_msa + strata(stratum)

other_fml_unadj <- als ~ other * lives_in_msa + strata(stratum)

other_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   other * lives_in_msa + strata(stratum)

falls_fml_unadj <- als ~ falls * lives_in_msa + strata(stratum)

falls_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   falls * lives_in_msa + strata(stratum)
```

Sixth, does the risk increase with each additional symptom?

```{r}
number_fml_unadj <- als ~ as.factor(n_total) * lives_in_msa + strata(stratum)

number_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   as.factor(n_total) * lives_in_msa + strata(stratum)
```


Quick note: all of these models omit year, age, and sex. This is because the
data are matched on age, sex, and enrollment start/end dates. Including the 
strata accounts for these factors while also properly clustering the errors.

# Estimation
## Does Having Any Code Increase the Risk?
Without adjustment:

```{r}
clogit(
  any_fml_unadj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

The odds are increased following one or more symptom (OR = 7.44), for those 
living in MSAs (OR = 1.43), but there is a multiplicative decrease in odds for
ALS for those with symptoms living in an MSA (OR = 0.84). The odds are still
increased (OR = $0.84 * 7.44 = 6.25$) but less than for those who do not live
in MSAs.

After adjustment, the story stays the same:

```{r}
clogit(
  any_fml_fulladj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

## Are There Differences Between Bulbar and Motor Symptoms?
Bulbar symptoms alone increases the odds before adjustment:

```{r}
clogit(
  any_bulbar_unadj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

And unsurprisingly adjustment does not change the story:

```{r}
clogit(
  any_bulbar_fulladj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

The same is true of motor symptoms

```{r}
clogit(
  any_motor_unadj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

and with adjustment:

```{r}
clogit(
  any_motor_fulladj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

## Does the Risk Increase With Symptoms From Different Origins?
Without adjustment:

```{r}
clogit(
  number_origin_unadj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

And with adjustment

```{r}
clogit(
  number_origin_fulladj,
  model_data
) |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

## By Symptom
### Bulbar Symptoms
#### Speech Problems

```{r}
speech_unadjusted <- clogit(
  speech_fml_unadj,
  model_data
)

speech_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
speech_adjusted <- clogit(
  speech_fml_fulladj,
  model_data
)

speech_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

#### Swallowing Problems

```{r}
swallowing_unadjusted <- clogit(
  swallowing_fml_unadj,
  model_data
)

swallowing_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
swallowing_adjusted <- clogit(
  swallowing_fml_fulladj,
  model_data
)

swallowing_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

### Motor Problems
#### Strength Problems
```{r}
strength_unadjusted <- clogit(
  strength_fml_unadj,
  model_data
)

strength_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
strength_adjusted <- clogit(
  strength_fml_fulladj,
  model_data
)

strength_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

#### Gait Problems
```{r}
gait_unadjusted <- clogit(
  gait_fml_unadj,
  model_data
)

gait_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
gait_adjusted <- clogit(
  gait_fml_fulladj,
  model_data
)

gait_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

#### Involuntary Movements
```{r}
involuntary_movements_unadjusted <- clogit(
  involuntary_movement_fml_unadj,
  model_data
)

involuntary_movements_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
involuntary_movements_adjusted <- clogit(
  involuntary_movement_fml_fulladj,
  model_data
)

involuntary_movements_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

#### Pain
```{r}
pain_unadjusted <- clogit(
  pain_fml_unadj,
  model_data
)

pain_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
pain_adjusted <- clogit(
  pain_fml_fulladj,
  model_data
)

pain_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

#### Falls
```{r}
falls_unadjusted <- clogit(
  falls_fml_unadj,
  model_data
)

falls_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
falls_adjusted <- clogit(
  falls_fml_fulladj,
  model_data
)

falls_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

#### Other
```{r}
other_unadjusted <- clogit(
  other_fml_unadj,
  model_data
)

other_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
other_adjusted <- clogit(
  other_fml_fulladj,
  model_data
) 

other_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

To summarize these in a practical way:

```{r}
list(
  speech_unadjusted,
  swallowing_unadjusted,
  strength_unadjusted,
  gait_unadjusted,
  involuntary_movements_unadjusted,
  pain_unadjusted,
  falls_unadjusted,
  other_unadjusted
) |>
  lapply(
    broom::tidy,
    exponentiate = TRUE,
    conf.int = TRUE
  ) |>
  enframe() |>
  mutate(
    symptom = c("Speech", "Swallowing", "Muscle",
                "Gait", "Involuntary\nMovements", "Pain", "Falls",
                "Other Motor")
  ) |>
  unnest(cols = value) |>
  mutate(
    variable = rep(c("Symptom", "Lives in Urban Area", "Symptom x Lives in Urban Area"), 8)
  ) |>
  mutate(
    symptom = forcats::fct_relevel(
      symptom,
      "Speech", "Swallowing", "Muscle",
      "Gait", "Involuntary\nMovements", "Pain", "Falls",
      "Other Motor"
    ),
    variable = forcats::fct_relevel(
      variable,
      "Symptom", "Lives in Urban Area", "Symptom x Lives in Urban Area"
    )
  ) |>
  ggplot(
    aes(
      x = symptom,
      y = estimate,
      ymin = conf.low,
      ymax = conf.high
    )
  ) + 
  geom_point() +
  geom_linerange() +
  facet_grid(rows = vars(variable), scale = "free_y") +
  geom_hline(yintercept = 1, linetype = 2) +
  labs(x = "", y = "Odds Ratio\n(95% CI)") +
  theme_bw()
```

```{r}
list(
  speech_adjusted,
  swallowing_adjusted,
  strength_adjusted,
  gait_adjusted,
  involuntary_movements_adjusted,
  pain_adjusted,
  falls_adjusted,
  other_adjusted
) |>
  lapply(
    broom::tidy,
    exponentiate = TRUE,
    conf.int = TRUE
  ) |>
  enframe() |>
  mutate(
    symptom = c("Speech", "Swallowing", "Muscle",
                "Gait", "Involuntary\nMovements", "Pain", "Falls",
                "Other Motor")
  ) |>
  unnest(cols = value) |>
  group_by(symptom) |>
  filter(row_number() >= 35) |>
  ungroup() |>
  mutate(
    variable = rep(c("Symptom", "Lives in Urban Area", "Symptom x Lives in Urban Area"), 8)
  ) |>
  mutate(
    symptom = forcats::fct_relevel(
      symptom,
      "Speech", "Swallowing", "Muscle",
      "Gait", "Involuntary\nMovements", "Pain", "Falls",
      "Other Motor"
    ),
    variable = forcats::fct_relevel(
      variable,
      "Symptom", "Lives in Urban Area", "Symptom x Lives in Urban Area"
    )
  ) |>
  ggplot(
    aes(
      x = symptom,
      y = estimate,
      ymin = conf.low,
      ymax = conf.high
    )
  ) + 
  geom_point() +
  geom_linerange() +
  facet_grid(rows = vars(variable), scale = "free_y") +
  geom_hline(yintercept = 1, linetype = 2) +
  labs(x = "", y = "Odds Ratio\n(95% CI)") +
  theme_bw() + 
  scale_y_log10()
ggsave("~/projects/als-sx/fig_1.svg", width = 8, height = 8)
```

### Number of Symptoms
```{r}
n_symptoms_unadjusted <- clogit(
  number_fml_unadj,
  model_data
)

n_symptoms_unadjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)
```

```{r}
n_symptoms_adjusted <- clogit(
  number_fml_fulladj,
  model_data
)

n_symptoms_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  knitr::kable(digits = 2)

n_symptoms_adjusted |> 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
  filter(!stringr::str_detect(term, "msa"), stringr::str_detect(term, "as.factor")) |>
  mutate(
    variable = 1:8
  ) |>
  ggplot(aes(x = variable, 
             y = estimate,
             ymin = conf.low,
             ymax = conf.high)) + 
  geom_point() + 
  geom_linerange() + 
  labs(x = "Number of Symptoms", y = "ORs Compared to Having No Symptoms\n(95% CI)") + 
  theme_bw()
```

## Excluding Recent "Exposures"
Require the exposure to have occurred at least $i$ days before the ALS diagnosis
or data exit date where $i$ ranges from 0 days to 5 years. 

Since this difference is only defined for people who had a symptom, retain any
people who are symptom free if they're possibly still included in the data 
(e.g., if the time between the enrollment date and the end date is less than 
$i$ days). 

First, a couple specific models:

```{r}
delay_30 <- clogit(
  any_fml_fulladj,
  model_data |>
    filter(is.na(first_symptom_date) | (first_symptom_date < (end_date - 30))) |>
    filter(first_date < (end_date - 30)) |>
    mutate(
      across(
        is_logical,
        as.numeric
      )
    )
  )

delay_60 <- clogit(
  any_fml_fulladj,
  model_data |>
    filter(is.na(first_symptom_date) | (first_symptom_date < (end_date - 60))) |>
    filter(first_date < (end_date - 60)) |>
    mutate(
      across(
        is_logical,
        as.numeric
      )
    )
  )

delay_90 <- clogit(
  any_fml_fulladj,
  model_data |>
    filter(is.na(first_symptom_date) | (first_symptom_date < (end_date - 90))) |>
    filter(first_date < (end_date - 90)) |>
    mutate(
      across(
        is_logical,
        as.numeric
      )
    )
  )

delay_180 <- clogit(
  any_fml_fulladj,
  model_data |>
    filter(is.na(first_symptom_date) | (first_symptom_date < (end_date - 180))) |>
    filter(first_date < (end_date - 180)) |>
    mutate(
      across(
        is_logical,
        as.numeric
      )
    )
  )

delay_270 <- clogit(
  any_fml_fulladj,
  model_data |>
    filter(is.na(first_symptom_date) | (first_symptom_date < (end_date - 270))) |>
    filter(first_date < (end_date - 270)) |>
    mutate(
      across(
        is_logical,
        as.numeric
      )
    )
  )

delay_365 <- clogit(
  any_fml_fulladj,
  model_data |>
    filter(is.na(first_symptom_date) | (first_symptom_date < (end_date - 365))) |>
    filter(first_date < (end_date - 365)) |>
    mutate(
      across(
        is_logical,
        as.numeric
      )
    )
  )

glance_vars <- c("nobs","nevent", "concordance", "r.squared")
tbl_merge(
  tbls = list(
    tbl_regression(delay_30, exponentiate = TRUE) |> add_glance_table(glance_vars),
    tbl_regression(delay_60, exponentiate = TRUE) |> add_glance_table(glance_vars),
    tbl_regression(delay_90, exponentiate = TRUE) |> add_glance_table(glance_vars),
    tbl_regression(delay_180, exponentiate = TRUE) |> add_glance_table(glance_vars),
    tbl_regression(delay_270, exponentiate = TRUE) |> add_glance_table(glance_vars),
    tbl_regression(delay_365, exponentiate = TRUE) |> add_glance_table(glance_vars)
  ),
  tab_spanner = c("**30 Day Delay**", "**60 Day Delay**", "**90 Day Delay**", 
                  "**180 Day Delay**", "**270 Day Delay**", "**365 Day Delay**")
)
```

And then with all the lags up to 5 years:

```{r}
lags <- round(seq(0, 2, by = 1/12) * 365)
results <- vector("list", length(lags))
j <- 1
p <- progress::progress_bar$new(
  total = length(lags), 
  format = "[:bar] :percent in :elapsed with :eta remaining"
)
for (i in lags) {
  results[[j]] <- clogit(
    any_fml_fulladj,
    model_data |>
      filter(is.na(first_symptom_date) | (first_symptom_date < (end_date - i))) |>
      filter(first_date < (end_date - i))
    ) |> 
      broom::tidy(exponentiate = TRUE, conf.int = TRUE) |>
      mutate(i = i) |>
      select(i, term, estimate, conf.low, conf.high)
    j <- j + 1
    p$tick()
}

results |>
  bind_rows() |> 
  filter(
    stringr::str_detect(term, "Neuro")
  ) |>
  ggplot(aes(x = i / 365, 
             y = estimate,
             ymin = conf.low,
             ymax = conf.high)) + 
  geom_point() +
  geom_linerange() + 
  labs(x = "Lead-in Period by Years",
       y = "Odds Ratio for Elixhauser Neurological CMs") + 
  theme_bw()
ggsave("~/projects/als-sx/neuro_series.svg", width = 16/2, height = 9/2)

results |>
  bind_rows() |> 
  filter(
    stringr::str_detect(term, "sx") | stringr::str_detect(term, "msa")
  ) |>
  mutate(
    term = case_when(
      term == "any_sxTRUE" ~ "Has Symptom",
      term == "lives_in_msaTRUE" ~ "Lives in MSA",
      term == "any_sxTRUE:lives_in_msaTRUE" ~ "Symptom x MSA"
    ),
    term = forcats::fct_relevel(term,
      "Has Symptom", "Lives in MSA")
  ) |>
  ggplot(
    aes(
      x = i / 365,
      y = estimate,
      ymin = conf.low,
      ymax = conf.high
    )
  ) + 
  geom_point() + 
  geom_linerange() + 
  facet_grid(rows = vars(term), scales = "free_y") + 
  labs(x = "Years Between Symptom and ALS Date", y = "Odds Ratio") +
  theme_bw()
ggsave("~/projects/als-sx/symptom_series.svg", width = 16/2, height = 9/2)
```