---
title: Main Analysis
author: Jacob Simmering
format: 
  html:
    embed-resources: true
---

```{r pkg-load}
library(tidyverse)
library(survival)
library(tableone)
```

# Data Loading and Processing
```{r}
model_data <- read_rds("/Shared/lss_jsimmeri/als/model_data.rds")
```

```{r}
model_data <- model_data |>
  mutate(
    age = first_year - dobyr,
    female = sex == 1,
    stratum = as.numeric(as.factor(glue::glue("{first_year}-{last_year}-{dobyr}-{sex}")))
  ) |>
  mutate(n_sx = fall + op + muscle,
         any_sx = fall | op | muscle)
```

# Data Summary
First, the data summary by ALS status:

```{r}
all_vars <- c(
  "age", "female", 
  "first_year", "last_year",
  "rate_outpatient_visits", "intensity_outpatient_visits_dx",
  "rate_inpatient_stays", "intensity_inpatient_stays_los",
  "elix_CHF", "elix_Valvular", "elix_PHTN", "elix_PVD", "elix_HTN",
  "elix_HTNcx", "elix_Paralysis", "elix_NeuroOther", "elix_Pulmonary", "elix_DM",
  "elix_DMcx", "elix_Hypothyroid", "elix_Renal", "elix_Liver", "elix_PUD", 
  "elix_HIV", "elix_Lymphoma", "elix_Mets", "elix_Tumor", "elix_Rheumatic", 
  "elix_Coagulopathy", "elix_Obesity", "elix_WeightLoss", "elix_FluidsLytes", "elix_BloodLoss", 
  "elix_Anemia", "elix_Alcohol", "elix_Drugs", "elix_Psychoses", "elix_Depression", 
  "lives_in_msa",
  "any_sx", "n_sx", 
  "fall", "muscle", "op")
factor_vars <- c("n_sx")
nonnormal_vars <- c("age", "first_year", "last_year", "rate_outpatient_visits", 
  "intensity_outpatient_visits_dx", "rate_inpatient_stays", 
  "intensity_inpatient_stays_los")

CreateTableOne(
  data = model_data,
  vars = all_vars,
  factorVars = factor_vars,
  strata = c("als")
) |>
  print(nonnormal = nonnormal_vars, smd = TRUE)
```

And then by ALS and whether they live in an MSA:

```{r}
CreateTableOne(
  data = model_data,
  vars = all_vars,
  factorVars = factor_vars,
  strata = c("als", "lives_in_msa")
) |>
  print(nonnormal = nonnormal_vars, smd = TRUE)
```

# Define Models
I want to fit five models: 

First, does having any of these codes increase the odds of getting an ALS 
diagnosis?

```{r}
any_fml_unadj <- als ~ I(fall | muscle | op) * lives_in_msa + strata(stratum)

any_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   I(fall | muscle | op) * lives_in_msa + strata(stratum)
```

Second, does the risk increase with each additional code?

```{r}
number_fml_unadj <- als ~ as.factor(I(fall + muscle + op)) * lives_in_msa + strata(stratum)

number_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   as.factor(I(fall + muscle + op)) * lives_in_msa + strata(stratum)
```

And for models three to five, does the risk vary between the codes?

```{r}
fall_fml_unadj <- als ~ fall * lives_in_msa + strata(stratum)

fall_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   fall * lives_in_msa + strata(stratum)

muscle_fml_unadj <- als ~ muscle * lives_in_msa + strata(stratum)

muscle_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   muscle * lives_in_msa + strata(stratum)

op_fml_unadj <- als ~ op * lives_in_msa + strata(stratum)

op_fml_fulladj <- als ~ rate_outpatient_visits + intensity_outpatient_visits_dx + 
                   rate_inpatient_stays + intensity_inpatient_stays_los +
                   elix_CHF + elix_Valvular + elix_PHTN + elix_PVD + elix_HTN +
                   elix_HTNcx + elix_Paralysis + elix_NeuroOther + elix_Pulmonary + elix_DM +
                   elix_DMcx + elix_Hypothyroid + elix_Renal + elix_Liver + elix_PUD + 
                   elix_HIV + elix_Lymphoma + elix_Mets + elix_Tumor + elix_Rheumatic + 
                   elix_Coagulopathy + elix_Obesity + elix_WeightLoss + elix_FluidsLytes + elix_BloodLoss + 
                   elix_Anemia + elix_Alcohol + elix_Drugs + elix_Psychoses + elix_Depression + 
                   op * lives_in_msa + strata(stratum)
```

Quick note: all of these models omit year, age, and sex. This is because the
data are matched on age, sex, and enrollment start/end dates. Including the 
strata accounts for these factors while also properly clustering the errors.

# Estimation
## Does Having Any Code Increase the Risk?
Without adjustment:

```{r}
clogit(
  any_fml_unadj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

The odds are increased following one or more symptom (OR = 8.95), for those 
living in MSAs (OR = 1.49), but there is a multiplicative decrease in odds for
ALS for those with symptoms living in an MSA (OR = 0.79). The odds are still
increased (OR = $0.79 * 8.95 = 7.05$) but less than for those who do not live
in MSAs.

After adjustment, the story stays the same:

```{r}
clogit(
  any_fml_fulladj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

The odds are remain increased following one or more symptom (OR = 6.40), for 
those living in MSAs (OR = 1.47), but there is a multiplicative decrease in odds
for ALS for those with symptoms living in an MSA (OR = 0.79). The odds are still
increased (OR = $0.79 * 6.40 = 5.06$) but less than for those who do not live
in MSAs.

## Does the Risk Increase With More Risk Factors?
Without adjustment:

```{r}
clogit(
  number_fml_unadj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

There is a dose-response pattern for more symptoms: OR = 6.05, 19.26, and 30.94. 
MSA remains a risk factor at 1.40. The interaction between MSA and the symptoms
does not vary by the number of symptoms (0.80, 0.70, 0.75). 

```{r}
clogit(
  number_fml_fulladj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

The dose response pattern is robust to adjustment: 5.09, 13.09, 16.30. The 
MSA effect is similar to before (OR = 1.48) and again does not vary with the 
MSA status (0.80, 0.73, 0.87). 

## By Risk Factor
### Falls and Injuries
Without adjustment:

```{r}
clogit(
  fall_fml_unadj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

With adjustment:

```{r}
clogit(
  fall_fml_fulladj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

### Muscle Problems
Without adjustment:

```{r}
clogit(
  muscle_fml_unadj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

With adjustment:

```{r}
clogit(
  muscle_fml_fulladj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

### Muscle Problems
Without adjustment:

```{r}
clogit(
  op_fml_unadj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```

With adjustment:

```{r}
clogit(
  op_fml_fulladj,
  model_data
) |> 
  broom::tidy() |>
  mutate(
    or = exp(estimate),
    lb = exp(estimate - 1.96 * std.error),
    ub = exp(estimate + 1.96 * std.error)
  ) |>
  select(term, or, lb, ub) |>
  knitr::kable(digits = 2)
```
